---
title: "TGC Tournament Winds"
author: "fuzion"
date: "September 7th 2015"
output:  
     html_document:  
       keep_md: true
---
### Introduction
I have collected a sample of wind speeds and directions in tournaments for [The Golf Club](http://www.thegolfclubgame.com). 
These are both from rounds that I have played, and from other people (usually by watching their rounds on [twitch](http://www.twitch.tv/directory/game/The%20Golf%20Club)).

I am using [R](https://www.r-project.org/) for the data analysis and presentation.
I include the data in full (at the end of this document), and the code used to generate this analysis in order to make it reproducable.

### Data import

```{r load_packages, include=FALSE}
library(broman)
library(formatR)
```
```{r message=FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
```

First, I read in the data.
```{r}
url <- "C:/Users/Matt/Documents/speedgolf/windspeeds.txt" # Point this to the data.
data <- read.table(url, header = TRUE, stringsAsFactors = FALSE)
head(data)
```
The column names show the (abbreviated) name of the tournament, the theme of the course, the round number, the wind speed (mph) on the first tee, the wind direction as given by the cardinal directions, and whether the tournament is a tour or a tournament.
```{r}
# Convert cardinal direction to angle (degrees)
cardinal_conv <- data.frame(direction = c("N", "NE", "E", "SE", "S", "SW", "W", "NW"), angle = c(0, 45, 90, 135, 180, 225, 270, 315), stringsAsFactors = FALSE)
data %<>% left_join(cardinal_conv, by = "direction")
```
### Analysis
We can calculate some statistics for the wind speeds.
This can also be split by theme.
```{r}
overview <- data %>% summarise(n = n(), mean_speed = mean(speed), sd_speed = sd(speed), err_mean = sd_speed/sqrt(n))
overview

overview_theme <- data %>% group_by(theme) %>% summarise(n = n(), mean_speed = mean(speed), sd_speed = sd(speed), err_mean = sd_speed/sqrt(n))
overview_theme
```
**We can see that the average wind speed is `r myround(overview$mean_speed, 1)` &plusmn; `r myround(overview$err_mean, 1)` mph.**

### Pretty plots!
OK, here's the part everyone was waiting for...
```{r}
wind_qs <- quantile(data$speed, c(0.25,0.5,0.75))
wind_histo <- ggplot(data = data, aes(x = speed)) +
    geom_histogram(aes(y = ..density..), binwidth = 2, fill = "black", colour = "gray", alpha = 0.1) +
    geom_density(alpha = 0.1, fill = "#FF0022", from = 0, to = 22, size = 1) +
    geom_vline(xintercept = wind_qs, colour = "red", linetype = "dotted", size = 0.8) +
    geom_vline(aes(xintercept = mean(speed)), colour = "blue", linetype = "dashed", size = 1) +
    scale_x_continuous(lim = c(0,24), breaks = seq(0,24,4)) +
    labs(x = "Wind speed (mph)", y = "Probability density") +
    geom_text(data = data %>% summarise(n = n()),
              aes(label = paste0("n = ", n),
                  x = 18.5, y = 0.1), size = 7) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 14))
wind_histo
```

We can compare this to other typical distributions; here we have a normal cumulative distribution function (cdf; in blue) and one for a uniform distribution (in green).
```{r}
set.seed(1)
sample_size <- length(data$speed) * 100
uni_winds <- data.frame(speed = sample(1:22, size = sample_size, replace = TRUE))
norm_winds <- data.frame(speed = rnorm(sample_size, overview$mean_speed, overview$sd_speed)) %>% round()

wind_ecdf2 <- ggplot(data = data, aes(x = speed)) +
    stat_ecdf(geom = "area", alpha = 0.1, fill = "red", colour = "black") +
    #stat_ecdf(geom = "step", alpha = 0.1, colour = "black") +
    stat_ecdf(data = uni_winds, geom = "line", colour = "green", linetype = "dashed") +
    stat_ecdf(data = norm_winds, geom = "line", colour = "blue", linetype = "dashed") +
    coord_cartesian(xlim = c(0, 22)) +
    labs(x = "Wind speed (mph)", y = "Cumulative probability") +
    geom_text(data = data %>% summarise(n = n()),
              aes(label = paste0("n = ", n),
                  x = 3, y = 0.9), size = 7) +
    theme_bw() +
    theme(axis.text.x = element_text(size = 14))
wind_ecdf2
```

**We see that we only have a `r round((ecdf(data$speed)(9)) * 100, 1)`% chance to have single digit winds for a given tournament round.**

(A cumulative distribution gives you the probability of getting a value &le; x on a sample, so we read up from 9 mph and go left to find the chances of getting a wind speed of 9 or less.)

We can also look at the direction of the wind, and plot a form of windrose.

```{r}
windplot2 <- ggplot(data = data, aes(x = angle, y = speed)) +
    annotate("rect", xmin = 0, xmax = 360, ymin = 0, ymax = 5, alpha = 0.2, fill = "green") +
    annotate("rect", xmin = 0, xmax = 360, ymin = 5, ymax = 10, alpha = 0.2, fill = "blue") +
    annotate("rect", xmin = 0, xmax = 360, ymin = 10, ymax = 15, alpha = 0.2, fill = "yellow") +
    annotate("rect", xmin = 0, xmax = 360, ymin = 15, ymax = 22, alpha = 0.2, fill = "red") +
    geom_point(size = 6, shape = 21, colour = "black", aes(fill = theme)) +
    scale_x_continuous(breaks = seq(0,315,45), lim = c(0, 360), labels = cardinal_conv$direction) +
    scale_y_continuous(breaks = seq(0,24,5), lim = c(0, 22)) +
    labs(y = "Wind speed (mph)", fill = "Theme") +
    scale_fill_brewer(palette = "Set1") +
    coord_polar(theta = "x", start = 0, direction = 1) +
    theme_bw() +
    theme(axis.title.x = element_blank())
windplot2
```
```{r message=FALSE, results='hide', fig.show='hide'}
source("C:/Users/Matt/Documents/speedgolf/windrose.R")
windrose2 <- plot.windrose(data = data, spd = "speed", dir = "angle",
                   spdseq = seq(0,24,4),
                   spdmax = 22,
                   dirres = 45,
                   palette = "RdPu") +
    scale_x_discrete(labels = cardinal_conv$direction) +
    theme_bw() +
    theme(axis.title.x = element_blank())
```
```{r}
windrose2
```

### Summary

Tournament winds in The Golf Club tend to be slightly higher than [real life wind speeds](http://average-wind-speed.findthebest.com/) e.g. North Carolina: 7.5 mph; Chicago: 10.3 mph.
Also, they affect the ball significantly more than they should, even on 4 yard chip shots in game.
This makes the tournament side of the game intimidating for newer players.

Without more data, I cannot tell if the theme makes a significant difference to the wind speeds. 

I am in favour of more local gusts in-game, to better simulate the weather conditions, and also to limit the amount of exact calculation to offset the winds' effect.
Without winds, the game becomes very easy to calculate, e.g. see the 52 (-20) I shot yesterday in 1 mph winds in the HB ProAm tournament.

### Appendix
```{r}
data
```
